Testing the precision of Bangla Communication Scale: Item Response Theory Guided Analysis

### 2022-02-20 communication.Rmd
IRT description corrected
Repo is ready to go public

documentclass     : "apa6"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: false
    latex_engine: xelatex
    includes:
      in_header: header.tex
mainfont: Arial
---

@@

```{r data, include=F}
# data <- readxl::read_excel("Processeddata/communication_data_n_582.xlsx")
# readr::write_rds(data, "Processeddata/communication_data_n_582.rds")
data <- readr::read_rds("Processeddata/communication_data_n_582.rds")

@@

efa.data.full <- subset(data, Sample == "EFA")
cfa.data.full <- subset(data, Sample == "CFA"| Sample=="Validity")
validity.data.full <- subset(data, Sample == "Validity")

sem.data <- data[,c(2,6:66)]
efa.data <- subset(sem.data , Sample == "EFA")
efa.data <- efa.data[,c(2:24)]
efa.data <- na.omit(efa.data)

@@

irt.data <- sem.data[,2:24]
```

@@

# Study-1: Translations and Exploratory Factor Analysis
Study-1 had two objectives. First we translated the 23 items of the Communication Scale [@barkman2002four] from English to Bangla. Second, we conducted an Exploratory Factor Analysis (EFA) to identify the latent structure of the scale.

@@

```{r sescrptives, include=FALSE}
desc <- efa.data.full[,c(3,4,5)]
## Recoding desc$Sex
desc$Gender <- fct_recode(desc$Gender,
  "Boy" = "male",
  "Girl" = "female"
)
colnames(desc) <- c("Gender", "Age", "Socio Economic Status")

@@

@@

@@

We used R (version 4.1.0) [@R-base], including several R packages [@R-lavaan; @R-mirt; @R-psych; @R-tabledown], for our analyses. Prior to the data analysis normality assumptions were checked. Our data violeted both univariate and multivariate normality. As such we used polychoric correlation matrix and 'principal axis" factor extraction method to conduct the exploratory factor analysis [@watkinsStepbyStepGuideExploratory2020]. An oblique rotation method was employed. A posterior sampling adequacy was estimated using KMO statistics [@kaiserIndexFactorialSimplicity1974]. TO identify the optimum number of factors required to explain the latent structure of our scale we used Scree plot [@cattellScreeTestNumber1966a], Horn's parallel analysis [@hornRationaleTestNumber1965], Hull method [@lorenzo-sevaHullMethodSelecting2011] and Minimum average partials method (MAP) [@velicerDeterminingNumberComponents1976]. Lastly, to identify the simple structure, we followed the following guidelines  (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@childEssentialsFactorAnalysis2006a; @mulaikFoundationsFactorAnalysis2009; @watkinsStepbyStepGuideExploratory2020]

To estimate reliability we used internal consistency coefficient ordinal alpha. Aprt from being the most popular reliability coefficient,Cronbach's alpha coefficient has a tendency to deflate the estimates for  Likert type data as the calculation is based on pearson-correlation matrix which requires data with continuous measurement level [@gadermann2012estimating; @zumbo2007ordinal]. For better estimates of reliability we reported ordinal alpha using polychoric-correlation [@zumbo2007ordinal].

### Results & Discussion
```{r, gtfig, include =F}

@@

min.cor <- apa(min(abs(upper)),2,F) #minimum cor.coefficient of the matrix
max.cor <- apa(max(abs(upper)),2,F) # max. correlation coefficient of the matrix

@@

@@

corplot <- ggcorrplot(correlations$rho, hc.order = TRUE, outline.col = "white", type = "lower", 
           ggtheme = ggplot2::theme_minimal(),tl.srt = 90,colors = c("#E64B35FF", "white", "#3C5488FF") )+
  theme( panel.grid.major = element_blank(),
         axis.text.x = element_text(size = 8),
         axis.text.y = element_text(size = 8),
         panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))
#p.mat <- cor_to_p(correlations$rho, 428, method = "polychoric")
#to show pvalue in the fig add p.mat=p.mat
ggsave("Figures/corplot.png",corplot, width = 6, height = 6, dpi = 600, bg ="white")

@@

(ref:figCor) Inter item polychoric correlation coefficients for the 23 items. 9.8 % inter-item correlation coefficients were higher than |.30|

@@

```{r ParallelEFA, include=FALSE}
#This chunk holds code for factor identification method : Horn
# Parallel analysis
Horn <- paran(mat=correlations$rho, iterations=500, centile=0, quietly=FALSE,
           status=TRUE, all=FALSE, cfa=TRUE, graph=TRUE,
           color=TRUE, col=c("black","red","blue"),
           lty=c(1,2,3), lwd=1, legend=TRUE, 
           width=640, height=640, grdevice=png, seed=0, n=300)

@@

write.csv(full.map, "Table_raw/map_stat.csv")
```

@@

To identify optimum number of factor required to express the latent structure adequetly we at first used the Scree-plot [@cattellScreeTestNumber1966a]. The Scree-plot suggested one factor solution. One factor solution was also supported by MAP method [@velicerDeterminingNumberComponents1976] and Hull method [@lorenzo-sevaHullMethodSelecting2011]. Minimum average partial (MAP) method  expects the average squared off-diagonal values of the calculated partial correlation matrix to be minimum when the correct number of factors are extracted [@velicerDeterminingNumberComponents1976]. In our data set this value reached the minimum after extracting the first factor (Supplementary Table 2). Hull method tried to find an optimal number of factors to balance model fit and the number of parameters and offered one factor solution in our data set (Figure \@ref(fig:facIdFig)).

@@

factor <- cowplot::plot_grid(scree.plot,Hull, 
                   labels = "AUTO",
                   ncol=1, 
                  align = "v", 
                  label_fontfamily = "sans",
                  label_fontface = "plain")
ggsave("Figures/factors.png",factor, width = 5, height = 7, dpi = 600)
```

@@

knitr::include_graphics("Figures/factors.png", dpi =800)

@@

The initial solution of the fitted one factor model had six items with poor factor loadings (Item 1, 2,5,9, 16, 23). We discarded these items from the model and run another EFA. This iteration of EFA yielded a one factor simple structure with 17 items. The one factor solution explained `r apa(var3,2,T)`% of the total variance.  The internal consistency reliability coefficient for this model was, ordinal $\alpha$ = `r apa(one.factor.alpha,2,F)`.

@@

png(filename="figures/CFAplot.png", 
    type="cairo",
    units="in", 
    width=6, 
    height=6, 
    pointsize=12, 
    res=1200)
semPaths (fit.1 , 
          what= "std",
          #"hide", #(hides coefficeits)
          #whatLabels = "std",
          intercepts = F,
          style ="OpenMx",
          #residScale = 6,
          theme = "colorblind",
          nCharNodes = 0,
          reorder =T,
          rotation =2,
          layout ="tree",
          cardinal = T,
          curvePivot =T,
          sizeMan =8,#items length
          sizeMan2 = 2,#items
          sizeLat = 12,#factors
          thresholds = FALSE,
          equalizeManifests =F,
          fade = FALSE,
          edge.label.cex = .8,
          #exoCov = T,
          centerLevels = T,
          #edge.color="black",
          label.scale=T,
          label.cex=1.2, #Font of factor and item name
          residuals=FALSE,
           #exoVar=FALSE,
          #fixedStyle=6, #Style of arrow (guide item)
          #freeStyle=1#Style of arrow (other item), #XKCD = TRUE

@@

ggsave("Figures/personfit.jpeg",person , width = 12, height = 14, dpi = 600, bg = "white")

@@

ggsave("Figures/person.png",person , width = 12, height = 8, dpi = 600, bg = "white")

@@

We fitted BCS to the IRT framework using "graded response" model [@samejima1969estimation] to the combined sample of study 1 and study 2 (n= `r nrow(irt.data)`). It required 20 iterations (Log-likelihood -12974.970) for the  model to converge. For the fitted model, Among the `r nrow(items)` retained items the mean item discrimination was `r apa(mean(items$a),2,T)`± `r apa(sd(items$a),2,T)` (range = `r apa(min(items$a),2,T)`-`r apa(max(items$a),2,T)`). All 15 items had very high discrimination [@RN1275]. Table \@ref(tab:IRT-tab2) summarizes the item discrimination, item difficulty and item fit statistics of the revised models. All items has a moderate discrimination. Standardized outfit statistics ranged between `r apa(min(item.fit$z.outfit),2,T)`-`r apa(max(item.fit$z.outfit),2,T)` indicating a sufficient item fit to the models [@desjardinsHandbookEducationalMeasurement2018]. Sufficent item fit is also reflected by the RMSEA values associated with S-$\chi^2$ statistics ( `r apa(min(item.fit$RMSEA.S_X2),2,T)`-`r apa(max(item.fit$RMSEA.S_X2),2,T)`). Lastly, Fig\@ref(fig:person-fit) indicates that for the Zh is larger than -2 for most participants, suggesting a good person-fit.

@@

ggsave("Figures/ICC.png",ICC , width = 12, height =14, dpi = 600, bg = "white")

@@

test.info <- tabledown::ggtestinfo(irt.data.1, fit)+ apatheme+geom_line(colour = "black", size=1.5)+geom_area(fill ="#E64B3599" )
# +geom_vline(xintercept = -3, colour ="black")+
#   geom_vline(xintercept = .5, colour ="black")
ggsave("Figures/TIC.png",test.info , width = 6, height = 4, dpi = 600, bg = "white")

@@

```{r saveWrightmao,  warning=F, message=F}
jpeg("Figures/personmap.jpeg", width = 12, height = 8, quality = 100,units="in", res= 1080, pointsize = 12 )
library(WrightMap)
WrightMap::wrightMap(estimation, threshold , 
          item.prop = 0.65,
          main.title = "Person-Map",
          # dim.names = c("Victimization", "Aggression"),
          thr.sym.col.fg = itemlevelcolors,
          thr.sym.col.bg = itemlevelcolors,
          vertLines = TRUE,
          show.thr.lab = F,
          dim.color = brewer.pal(5, "Set3"),
          label.items.srt = 45,
          cex = 1.5,
          return.thresholds=F
          )
```

@@

ggsave("Figures/itemInfoplot.png",item.info.plot, width = 8, height =8, dpi = 1200, bg ="white")

@@

```{r conrel-print, include=F}
##
png(filename="Figures/conrel.png", 
    type="cairo",
    units="in", 
    width=12, 
    height=6, 
    pointsize=12, 
    res=300)
library(latticeExtra)
library(directlabels)
font.settings <- list(
  font = 1,
  cex = 6,
  fontfamily = "sans")

>>>>

documentclass     : "apa6"
classoption       : "man"
output            : 
  papaja::apa6_pdf:
    keep_tex: false
    latex_engine: xelatex
    includes:
      in_header: header.tex

mainfont: Arial
---

@@

```{r data, include=F}
# data <- readxl::read_excel("Processeddata/communication_data_n_622.xlsx")
# readr::write_rds(data, "Processeddata/communication_data_n_622.rds")
data <- readr::read_rds("communication_data_n_622.rds")

@@

efa.data.full <- subset(data, Sample == "EFA"| Sample=="Testretest")
cfa.data.full <- subset(data, Sample == "CFA"| Sample=="Validity")
validity.data.full <- subset(data, Sample == "Validity")

sem.data <- data[,c(2,6:66)]
efa.data <- subset(sem.data , Sample == "EFA"| Sample=="Testretest")
efa.data <- efa.data[,c(2:24)]
efa.data <- na.omit(efa.data)

@@

irt.data <- sem.data[,2:24]

testretest_data <- subset(data, Sample=="Testretest")
test <- testretest_data[, c(6:28)]
retest <- testretest_data[, c(67:89)]

all.item.test <- test %>% 
  rowwise() %>% 
  mutate(test_total=sum(across(contains("CS")), na.rm = T))

all.item.retest <- retest %>% 
  rowwise() %>% 
  mutate(retest_total=sum(across(contains("CSA")), na.rm = T))

all.item.icc <- as.data.frame(cbind(all.item.test$test_total, all.item.retest$retest_total))

#

# cor(all.item.test$test_total,all.item.retest$retest_total)

@@

# Study-1: Translations and Exploratory Factor Analysis

Study-1 had two objectives. First we translated the 23 items of the Communication Scale [@barkman2002four] from English to Bangla. Second, we conducted an Exploratory Factor Analysis (EFA) to identify the latent structure of the scale.

@@

```{r dsescrptives, include=FALSE}
desc <- efa.data.full[,c(3,4,5)]
## Recoding desc$Sex
desc$Gender <- fct_recode(desc$Gender,
  "Boy" = "male",
  "Girl" = "female"
)
colnames(desc) <- c("Gender", "Age", "Socio Economic Status")

@@

desc.test.retest.data <- testretest_data[,c(3,4,5)]

desc.test.retest.data$Gender <- fct_recode(desc.test.retest.data$Gender,
  "Boy" = "male",
  "Girl" = "female")
test.retest.description <- desc.test.retest.data %>% 
  psych::describeBy("Gender")
test.female <- as.tibble(test.retest.description$Girl)
test.male <- as.tibble(test.retest.description$Boy)

test.retest.des_tab <- desc.test.retest.data %>%
  tbl_summary(
    by = Gender,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    type = list(Age ~ 'continuous'),
    digits = all_continuous() ~ 2,

as_tibble(test.retest.des_tab) -> test.retset.desc_tibble

@@

To estimate the test-retest reliability we used intraclass correlation coefficients (ICC) . With 80% power and alpha at 0.05, a minimum sample size of 22 is recommended to detect the ICC value of .50 [@bujang2017simplified]. Considering potential instances of drop-out we randomly chose a subset of `r nrow(testretest_data)` participants from our study-1 sample. They responded to the Bangla communication scale twice, two week apart.  Among these `r nrow(testretest_data)` participants `r test.female[1,2]` were female aged between `r test.female[2,8]` to `r test.female[2,9]` years (`r round(test.female[2,3],2)`±`r round(test.female[2,4],2)`). `r test.male[1,2]`  were male with an age range between `r test.male[2,8]` participants `r test.male[2,9]` years (`r round(test.male[2,3],2)`±`r round(test.male[2,4],2)`). `r  test.retset.desc_tibble [4,2]` participants belonged to middle socio-economic status. `r  test.retset.desc_tibble [3,2]` and `r  test.retset.desc_tibble [5,2]` participants belonged to lower and upper socio-economic status-respectively.

@@

We used R (version 4.1.0) [@R-base], including several R packages [@R-lavaan; @R-mirt; @R-psych; @R-tabledown], At first we checked for missing values and outliers. To estimate test-retest reliability of the Bangla communication scale we used intraclass correlation coefficients computed based on single-rating, absolute-agreement, 2-way random-effects model. An ICC value higher than .50 is considered adequate [@koo2016guideline]. Our data violated both univariate and multivariate normality. As such we used polychoric correlation matrix as it is more robust towards these violations. For initial item analysis we calculated item-total correlation and estimated internal consistency reliability coefficient ordinal alpha drop statistics. Though Cronbach's alpha is the the most popular reliability coefficient,it has a tendency to deflate the estimates for  Likert type data as the calculation is based on pearson-correlation matrix which requires data with continuous measurement level [@gadermann2012estimating; @zumbo2007ordinal]. Since our data followed Likert types response format for better estimation of reliability we reported ordinal alpha using polychoric-correlation [@zumbo2007ordinal].

In the exploratory factor analysis we used 'principal axis" factor extraction method [@watkinsStepbyStepGuideExploratory2020]. A posterior sampling adequacy was estimated using KMO statistics [@kaiserIndexFactorialSimplicity1974]. TO identify the optimum number of factors required to explain the latent structure of our scale we used Scree plot [@cattellScreeTestNumber1966a], Horn's parallel analysis [@hornRationaleTestNumber1965], Hull method [@lorenzo-sevaHullMethodSelecting2011] and Minimum average partials method (MAP) [@velicerDeterminingNumberComponents1976]. Lastly, to identify the simple structure, we followed the following guidelines  (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@childEssentialsFactorAnalysis2006a; @mulaikFoundationsFactorAnalysis2009; @watkinsStepbyStepGuideExploratory2020]

```{r, gtfig, include =F}

@@

correlationtab <- as.data.frame(upper )

min.cor <- apa(min((upper)),2,F) #minimum cor.coefficient of the matrix
max.cor <- apa(max((upper)),2,F) # max. correlation coefficient of the matrix

@@

```{r, include = F}
cor.mat <- correlations$rho
cor.mat <- as.data.frame(cor.mat)
```

```{r results='asis'}
apa_table(cor.mat,landscape = T )
```

@@

corplot <- corplot <- ggcorrplot::ggcorrplot(correlations$rho, hc.order = TRUE, outline.col = "white", type = "lower", legend.title = "Correlation",
           ggtheme = ggplot2::theme_minimal(),tl.srt = 90,colors = c("#E64B35FF", "white", "#3C5488FF") )+
  theme( panel.grid.major = element_blank(),
         axis.text.x = element_text(size = 6),
         axis.text.y = element_text(size = 6),
         panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1),
         legend.text = element_text(size = 6),
         legend.title= element_text(size = 6))
#p.mat <- cor_to_p(correlations$rho, 428, method = "polychoric")
#to show pvalue in the fig add p.mat=p.mat
ggsave("Figures/corplot.png",corplot, width = 3.5, height = 3.5, dpi = 600, bg ="white")

@@

(ref:figCor) Inter item polychoric correlation coefficients for the 23 items. 11.56 % inter-item correlation coefficients were higher than |.30|

@@

```{r ParallelEFA, include=FALSE}
#This chunk holds code for factor identification method : Horn
# Parallel analysis
Horn <- paran(efa.data, iterations=500, centile=0, quietly=FALSE,
           status=TRUE, all=FALSE, cfa=TRUE, graph=TRUE,
           color=TRUE, col=c("black","red","blue"),
           lty=c(1,2,3), lwd=1, legend=TRUE, 
           width=640, height=640, grdevice=png, seed=0)

@@

```{r map}
papaja::apa_table(full.map, caption = "Minimum average partial (MAP) method of factor numder determination. MAP Statistics is the lowest in the 5th row indicating five factors are required.")
```

#### Test-retest reliability

```{r ICC}
ICC <- psych::ICC(all.item.icc)
papaja::apa_table(ICC$results, caption = "Intra Class Correlation.")
```

@@

To identify optimum number of factor required to express the latent structure adequately we at first used the Scree-plot [@cattellScreeTestNumber1966a]. The Scree-plot suggested one factor solution. One factor solution was also supported by MAP method [@velicerDeterminingNumberComponents1976] and Hull method [@lorenzo-sevaHullMethodSelecting2011]. Minimum average partial (MAP) method  expects the average squared off-diagonal values of the calculated partial correlation matrix to be minimum when the correct number of factors are extracted [@velicerDeterminingNumberComponents1976]. In our data set this value reached the minimum after extracting the first factor (Supplementary Table 2). Hull method tried to find an optimal number of factors to balance model fit and the number of parameters and offered one factor solution in our data set (Figure \@ref(fig:facIdFig)).

@@

factor <- cowplot::plot_grid(scree.plot,Hull, 
                   labels = "AUTO",
                   ncol=2, 
                  align = "h", 
                  label_fontfamily = "sans",
                  label_fontface = "plain")
ggsave("Figures/factors.png",factor, width = 7, height = 3.5, dpi = 600)
```

@@

knitr::include_graphics("Figures/factors.png", dpi =600)

@@

```{r 1f-tau-equivalance, eval=FALSE, include=FALSE}
tau.equivalance.2 <- coefficientalpha::tau.test(reduced.model.1F.1)
```
The initial solution of the fitted one factor model had six items with poor factor loadings (Item 1, 2,5,9, 16, 23). We discarded these items from the model and run another EFA. This iteration of EFA yielded a one factor simple structure with 17 items. The one factor solution explained `r apa(var3,2,T)`% of the total variance.  The internal consistency reliability coefficient for this model was, ordinal $\alpha$ = `r apa(one.factor.alpha,2,F)`.

@@

png(filename="Figures/CFAplot.png", 
    type="cairo",
    units="in", 
    width=3.5, 
    height=3.5, 
    pointsize=12, 
    res=600)
semPaths (fit.1 , 
          what= "std",
          #"hide", #(hides coefficeits)
          #whatLabels = "std",
          intercepts = F,
          style ="OpenMx",
          #residScale = 6,
          theme = "colorblind",
          nCharNodes = 0,
          reorder =T,
          rotation =2,
          layout ="tree",
          cardinal = T,
          curvePivot =T,
          sizeMan =8,#items length
          sizeMan2 = 2,#items
          sizeLat = 12,#factors
          thresholds = FALSE,
          equalizeManifests =F,
          fade = FALSE,
          edge.label.cex = .8,
          #exoCov = T,
          centerLevels = T,
          #edge.color="black",
          label.scale=T,
          label.cex=1.2, #Font of factor and item name
          residuals=FALSE,
           #exoVar=FALSE,
          #fixedStyle=6, #Style of arrow (guide item)
          #freeStyle=1#Style of arrow (other item), #XKCD = TRUE

@@

ggsave("Figures/personfit.jpeg",person , width = 3.5, height = 3.5, dpi = 600, bg = "white")

@@

ggsave("Figures/person.png",person , width = 3.5, height = 3.5, dpi = 600, bg = "white")

@@

We fitted BCS to the IRT framework using "graded response" model [@samejima1969estimation] to the combined sample of study 1 and study 2 (n= `r nrow(irt.data)`). It required 20 iterations (Log-likelihood -12974.970) for the  model to converge. For the fitted model, Among the `r nrow(items)` retained items the mean item discrimination was `r apa(mean(items$a),2,T)`± `r apa(sd(items$a),2,T)` (range = `r apa(min(items$a),2,T)`-`r apa(max(items$a),2,T)`).  Table \@ref(tab:irttab) summarizes the item discrimination, item difficulty and item fit statistics of the revised models. All items has a moderate discrimination. Standardized outfit statistics ranged between `r apa(min(item.fit$z.outfit),2,T)`-`r apa(max(item.fit$z.outfit),2,T)` indicating a sufficient item fit to the models [@desjardinsHandbookEducationalMeasurement2018]. Sufficent item fit is also reflected by the RMSEA values associated with S-$\chi^2$ statistics ( `r apa(min(item.fit$RMSEA.S_X2),2,T)`-`r apa(max(item.fit$RMSEA.S_X2),2,T)`). Lastly, Fig\@ref(fig:person-fit) indicates that for the Zh is larger than -2 for most participants, suggesting a good person-fit.

@@

ggsave("Figures/ICC.png",ICC , width = 7, height =7, dpi = 600, bg = "white")

@@

test.info <- tabledown::ggtestinfo(irt.data.1, fit)+ apatheme+geom_line(colour = "black", size=1.5)+geom_area(fill ="#E64B3599" )
# +geom_vline(xintercept = -3, colour ="black")+
#   geom_vline(xintercept = .5, colour ="black")
ggsave("Figures/TIC.png",test.info , width = 3.5, height = 3.5, dpi = 600, bg = "white")

@@

```{r saveWrightmao,  warning=F, message=F}
jpeg("Figures/personmap.jpeg", width = 7, height = 7, quality = 100,units="in", res= 1080, pointsize = 12 )
library(WrightMap)
WrightMap::wrightMap(estimation, threshold , 
          item.prop = 0.65,
          main.title = "Person-Map",
          # dim.names = c("Victimization", "Aggression"),
          thr.sym.col.fg = itemlevelcolors,
          thr.sym.col.bg = itemlevelcolors,
          vertLines = TRUE,
          show.thr.lab = F,
          dim.color = brewer.pal(5, "Set3"),
          label.items.srt = 45,
          cex = 1.5,
          return.thresholds=F
          )
```

@@

ggsave("Figures/itemInfoplot.png",item.info.plot, width = 7, height =7, dpi = 600, bg ="white")

@@

```{r conrel-print, include=F}
##
png(filename="Figures/conrel.png", 
    type="cairo",
    units="in", 
    width=3.5, 
    height=3.5, 
    pointsize=12, 
    res=600)
library(latticeExtra)
library(directlabels)
font.settings <- list(
  font = 1,
  cex = 6,
  fontfamily = "sans")
